An inside-outside algorithms for mutant RNAs with applications to error detections in structured RNAs

### 2012-10-06 Recomb/introduction_RECOMB.tex
intro

In this paper, we tackle the problem of exploring an rRNA sequence mutant space, to
identify positions which are probably sequencing error, given the  rRNA family sequences
and its consensus structure.

Leveraging the techniques  in \texttt{RNAmutants}~\cite{Waldispuhl2008}, and building on top 
of the \emph{Inside-Outside algorithm}, we define here a new method called \texttt{RNApyro}
efficiently computing for large RNAs those probabilities.
 Classical techniques define a probabilistic model using a Boltzmann distribution 
whose weights are based on the free energy of the structure, using as energy parameter
the values of Turner found in the NNDB~\cite{Turner2010}  for stacked, 
canonical and wobble, base pairs. As shown by Leontis and Westhof~\cite{Leontis2001},
this  does not encapsulated the large diversity of base pairs that any nucleotide
can form with any other, although with an energy too small to be yet determined
by experimental techniques. To quantify geometrical differences, they
 define an isostericity distance, increasing as two base pairs differ 
 more from one another in space. We incorporate this second measure in the Boltzmann weights.

To benchmark

>>>>

In this paper, we tackle the problem of, given an rRNA sequence its family and 
consensus secondary structure, 
identifing the positions which are probably sequencing error.

Leveraging the techniques  in \texttt{RNAmutants}~\cite{Waldispuhl2008}, and building on top 
of the \emph{Inside-Outside algorithm}, we define here a new method called \texttt{RNApyro}
efficiently computing for large RNAs those probabilities under a new 
pseudo-energetic model.
 Classical techniques define a probabilistic model using a Boltzmann distribution 
whose weights are based on the free energy of the structure, using as energy parameter
the values of Turner found in the NNDB~\cite{Turner2010}  for stacked, 
canonical and wobble, base pairs. As shown by Leontis and Westhof~\cite{Leontis2001},
this  does not encapsulated the large diversity of base pairs that any nucleotide
can form with any other, although with an energy too small to be yet determined
by experimental techniques. To quantify geometrical differences, they
 define an isostericity distance, increasing as two base pairs differ 
 more from one another in space. We incorporate this second measure in the Boltzmann weights.

To benchmark our technique we use the 5S ribosomal RNA family,
 which is around $120$ nucleotides longs. We show
